---
title: "Webscrpaing Regex in R"
author: "Shikha Chamoli"
urlcolor: blue
---

  > ### Objective
  >
  > * To use web scraping and regular expressions for analyzing a web page.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE
)
```

## 1) Version Control
```{r}
library(xml2)
library(rvest)
library(stringr)
library(ggplot2)
library(reshape2)
```

### 1a) Version history of stringr package
```{r}
readCranData <- function(packageName) {
  package_url = paste0("https://cran.r-project.org/src/contrib/Archive/",packageName,"/")
  download.file(package_url,'package_data.html')
  page_data <-read_html('package_data.html')
  allTables <-html_table(page_data)
  #print(length(allTables))
  data = allTables[[1]]
  data = head(tail(data[,2:5],-2),-1)
  data[is.na(data$Description),4] = ""
  data$`Last modified` = as.POSIXct(strptime(data$`Last modified`,format = "%Y-%m-%d %H:%M"))
  ## Parsing the size as numeric (in KB)
  data$Size = sapply(data$Size,function(str) {
    mult = switch(
      substr(str,nchar(str),nchar(str)),
      "K"=1.0,
      "M"=1024.0,
      "G"=1024.0*1024,
      1/1024.0)
    return(as.numeric(gsub("[K|M|G]","",str))*mult)
  })
  return(data)
}

stringrData = readCranData("stringr")
head(stringrData,3)
tail(stringrData,3)
```
### 1b) Step Plot of package size 
```{r}
plotPackageData <- function(packageName) {
  packageData = readCranData(packageName)
  plt<- ggplot() +
    geom_step(data=packageData, mapping=aes(x=`Last modified`, y=Size)) +
    geom_step(data=packageData, mapping=aes(x=`Last modified`, y=Size), direction="vh", linetype=3) +
    geom_point(data=packageData, mapping=aes(x=`Last modified`, y=Size), color="red") +
    labs(title=paste0("Package Size over time for package: ",packageName), subtitle = "Step Plot", y = "Package Size in KB", x = "Time")
  return(plt)
}

plotPackageData("stringr")
```

### 1c) Faceted grid plot for dplyr, ggplot2, knitr and stringr packages.
```{r}
packages=c("dplyr","ggplot2","knitr","stringr")
data = c()
for(i in 1:length(packages)){
  dx=readCranData(packages[i])
  dx$gridX = paste0("Package Size over time for package: ",packages[i])
  dx$gridY = ifelse(i<3,""," ")
  data = rbind(data,dx)
}

ggplot() +
  geom_step(data=data, mapping=aes(x=`Last modified`, y=Size)) +
  geom_step(data=data, mapping=aes(x=`Last modified`, y=Size), direction="vh", linetype=3) +
  geom_point(data=data, mapping=aes(x=`Last modified`, y=Size), color="red") +
  labs(title=paste0("Faceted plots for package size data "), subtitle = "Step Plot", y = "Package Size in KB", x = "Time") +
  facet_wrap(gridX ~ gridY) + 
  theme(strip.text.x = element_text(angle = 0,color = "blue",size = 10 ))

```

## 2) Package Meta Data
```{r}
getPackageData <- function(packageName){
  package_url = paste0("https://cran.r-project.org/web/packages/",packageName,"/index.html")
  download.file(package_url,'package_data.html')
  page_data <-read_html('package_data.html')
  allTables <-html_table(page_data)
  return(allTables)
}
```
### 2a) Count the number of external packages in the Imports field.
```{r}
getImports <- function(packageName){
  allTables=getPackageData(packageName)
  tab1 = allTables[[1]]
  imports = tab1$X2[tab1$X1 == "Imports:"]
  importList=strsplit(imports,",")[[1]]
  importList=gsub("[[:space:]]"," ", importList)
  return(importList)
}
length(getImports("dplyr"))
```
### 2b) Count the number of external packages in the Suggests field.
```{r}
getSuggests <- function(packageName){
  allTables=getPackageData(packageName)
  tab1 = allTables[[1]]
  suggests = tab1$X2[tab1$X1 == "Suggests:"]
  suggestsList=strsplit(suggests,",")[[1]]
  suggestsList=gsub("[[:space:]]"," ", suggestsList)
  return(suggestsList)
}
length(getSuggests("dplyr"))
```

### 2c) Count the number of authors (denoted by [aut]) in the Authorfield.
```{r}
getAuthors <- function(packageName){
  allTables=getPackageData(packageName)
  tab1 = allTables[[1]]
  authors = gsub("[[:space:]]"," ",tab1$X2[tab1$X1 == "Author:"])
  nameList = str_replace_all(authors, "\\s*\\[[^\\]]+\\]", "") # remove []
  tagList = str_trim(str_extract_all(authors, "\\s*\\[[^\\]]+\\]", ""))  # extract contents of []
  allAuthors=str_trim(str_split(nameList,",")[[1]])
  authorsList = allAuthors[str_detect(tagList,"aut")] # check for aut tag
  return(authorsList)
}
length(getAuthors("dplyr"))
```

### 2d) Count the number of package maintainers (denoted by [cre]) in the Author field.
```{r}
getMaintainers <- function(packageName){
  allTables=getPackageData(packageName)
  tab1 = allTables[[1]]
  authors = gsub("[[:space:]]"," ",tab1$X2[tab1$X1 == "Author:"])
  nameList = str_replace_all(authors, "\\s*\\[[^\\]]+\\]", "") # remove []
  tagList = str_trim(str_extract_all(authors, "\\s*\\[[^\\]]+\\]", ""))  # extract contents of []
  allAuthors=str_trim(str_split(nameList,",")[[1]])
  authorsList = allAuthors[str_detect(tagList,"cre")] # check for cre tag
  return(authorsList)}
length(getMaintainers("dplyr"))
```

### 2e) Count the total number of packages in the Reverse depends field.
```{r}
getReverseDepends <- function(packageName) {
  allTables=getPackageData(packageName)
  tab1 = allTables[[3]]
  depends = tab1$X2[gsub("[[:space:]]", " ", tab1$X1[1]) == "Reverse depends:"]
  dependsList=strsplit(depends,",")[[1]]
  return(dependsList)
}
length(getReverseDepends("dplyr"))
```
## 3) The Tidyverse: Cursory Exploration
### 3a) Print "version number" and "date published" in human readable format.
```{r}
printPackageVersion <- function(packageName) {
  tab1 = getPackageData(packageName)[[1]]
  print(paste0("The ",
             packageName,
             " package (version number ",
             tab1$X2[tab1$X1 == "Version:"][[1]],
             ") was published on ",
             tab1$X2[tab1$X1 == "Published:"][[1]]))
}
printPackageVersion("tidyverse")
```
### 3b) Print "Author" in human readable format.
```{r}
printAuthorInfo <- function(packageName) {
  authorsList= getAuthors(packageName)
  authorStr = ifelse(
    length(authorsList)<=1,
    authors,
    paste0(
      paste(authorsList[1:length(authorsList)-1], collapse = ", "), 
      " and ",
      authorsList[length(authorsList)]))
  print(paste0("The ",packageName," package was written by ",authorStr))
}

printAuthorInfo("tidyverse")
```
### 3c) List name of packages in Imports field.
```{r}
tidyverseImports = getImports("tidyverse")
```
### 3d) Identify the Tidyverse Import with the smallest acceptable version number
```{r}
minVersion = str_extract(tidyverseImports,regex("[\\d.]+"))
packageNames = str_trim(str_extract(tidyverseImports,regex("[^\\(]+")))
versionData = cbind(packageNames,minVersion)
sortedData = versionData[order(minVersion, na.last = TRUE),]
sortedData
print(paste0("The package ",sortedData[1,1], " has the smallest min acceptable version (",sortedData[1,2],")."))
```
### 3e) Count the number of Tidyverse Imports whose names end in the letter r.
```{r}
sum(str_detect(packageNames, regex("r$")))
```
### 3f) List the Tidyverse Imports whose names do not comprise solely lower-case letters.
```{r}
packageNames[str_detect(packageNames, regex("[^(a-z)]"))]
```

## 4) The Tidyverse: Deep Dive

### 4a) Write a function to fetch a property from summary table of any package, on it CRAN page.
```{r}
getSummaryProperty <- function(packageName,propertyName){
  allTables=getPackageData(packageName)
  tab1 = allTables[[1]]
  propertyValue = tab1$X2[startsWith(tab1$X1,propertyName)]
  if(length(propertyValue)<1) {
    return(c())
  }
  propertyStr = gsub("[[:space:]]"," ",propertyValue)
  cleanStr = str_replace_all(propertyStr, "\\s*\\([^\\)]+\\)", "") #remove ()
  cleanStr = str_replace_all(cleanStr, "\\s*\\[[^\\]]+\\]", "") # remove []
  cleanStr = str_replace_all(cleanStr, "\\s*\\<[^\\>]+\\>", "") # remove <>
  propertyList=str_trim(str_split(cleanStr,",")[[1]])
  return(propertyList)
}
getSummaryProperty("tidyverse", "Version")
getSummaryProperty("tidyverse", "Depends")
getSummaryProperty("tidyverse", "Imports")
getSummaryProperty("tidyverse", "Suggests")
getSummaryProperty("tidyverse", "Published")
getSummaryProperty("tidyverse", "Author")
getSummaryProperty("tidyverse", "Maintainer")
getSummaryProperty("tidyverse", "URL")
```

### 4b) Create and adjacency matrix
```{r}
tidyverseImports = getSummaryProperty("tidyverse", "Imports")
df <- data.frame(matrix(ncol = length(tidyverseImports), nrow = 0))
colnames(df) <- tidyverseImports

for(package in tidyverseImports){
  packageImports = getSummaryProperty(package, "Imports")
  if(length(packageImports)>0) {
    df[packageImports,package]=1
  }
}

df[is.na(df)]=0
rows=dim(df)[1]
cols=dim(df)[2]
df[1:7,1:7]
df[(rows-7):rows,(cols-7):cols]
```

### 4c) Plot histogram of number of imports of each tidyverse import.
```{r}
histData=melt(df)
histData= histData[histData$value>0,]
colnames(histData) = c("PackageName", "Value")
ggplot(histData, aes(x=PackageName)) + 
  geom_histogram(stat = "count",color="black", fill="lightblue")+
  labs(title="Histogram: Number of imports Per Tydiverse import", x= "Package Name", y = "Number of Imports") + 
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
```

### 4d) List the names of any Tidyverse Imports that require compilation from source files.
```{r}
compiledPackages = c()
for(package in tidyverseImports){
  needsCompilation = getSummaryProperty(package, "NeedsCompilation")
  if(needsCompilation=="yes") {
    compiledPackages= c(compiledPackages, package)
  }
}
print(compiledPackages)
```


